Lesson 3
========================================================

***

### What to Do First?
Notes:

***

### Pseudo-Facebook User Data
Notes: our goal is to understand user behaviour and their demographics.
What they are doing on Facebook and what they use.

```{r Pseudo-Facebook User Data}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)

names(pf)
qplot(x = dob_day, data = pf) +
  scale_x_discrete(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)
```

***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data = pf, x = friend_count)
```

#### How is this plot similar to Moira's first plot?
Response:

*** Most of the data are in the left part of the graph. Exponential shape

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(data = pf, x = friend_count, xlim = c(0, 1000))

qplot(data = pf, x = friend_count) +
  scale_x_continuous(limits = c(0, 1000))

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```
In the alternate solution below, the period or dot in the formula for facet_grid() represents all of the other variables in the data set. Essentially, this notation splits up the data by gender and produces three histograms, each having their own row. 
```{r}
    qplot(x = friend_count, data = pf) + 
      facet_grid(gender ~ .) 
```
Equivalent ggplot syntax: 
```{r}
    ggplot(aes(x = friend_count), data = pf) + 
      geom_histogram() + 
      scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
      facet_wrap(~gender)
```      
***

### Omitting NA Values
Notes: R adds a separate histogram for missing values of the grouping variable gender.

```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

***

### Statistics 'by' Gender
Notes: We want summary of friend count by gender

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes: Explore tenure - how many days someone has used facebook

```{r Tenure}
qplot(x = tenure, data = pf, binwidth = 30,
      color = I('black'), fill = I('#099DD9'))
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
    qplot(x = tenure/365, data = pf, binwidth = 0.25,
          color = I('black'), fill = I('#F79420')) +
          scale_x_continuous(limits = c(0, 7), breaks = seq(1, 7, 1))
```

***

### Labeling Plots
Notes: add xlab = '...' and ylab = '...' parameters

```{r Labeling Plots}
    qplot(x = tenure/365, data = pf, binwidth = 0.25,
          xlab = 'Number of years using facebook',
          ylab = 'Number of users in sample',
          color = I('black'), fill = I('#F79420')) +
          scale_x_continuous(limits = c(0, 7), breaks = seq(1, 7, 1))
```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf,
          color = I('red'), fill = I('#F79420')) +
  scale_x_discrete(breaks = seq(0,140,10))

qplot(x = age, data = pf, binwidth = 5,
          xlab = 'Age',
          ylab = 'Number of Facebook users',
          color = I('red'), fill = I('#F79420')) +
  scale_x_continuous(breaks = seq(0,140,10))
```

#### What do you notice?
Response: Peak for young people at about 17-25 years, plateau for middle age users than gradual decrease. Strange peak for ages above 90. There are regular increases at 23, 33, 43,53, 63, 73, 83, 93 and 103 years

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:
```
    summary(pf$friend_count)
    summary(log10(pf$friend_count + 1))
    summary(sqrt(pf$friend_count))
```
***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

install.packages("gridExtra")
library(gridExtra)

p0 = qplot(x = friend_count, data = pf, binwidth = 0.1,
               xlab = 'Number of friends',
               ylab = 'Number of Facebook users',
               color = I('red'), fill = I('#F79420')) +
scale_x_continuous(breaks = seq(0,5000,1000))

p1 = qplot(x = friend_count, data = pf, binwidth = 0.1,
               xlab = 'Number of friends',
               ylab = 'Number of Facebook users',
               color = I('red'), fill = I('#F79420')) +
scale_x_log10(breaks = c(0.1, 1, 10, 100, 1000))

p2 = qplot(x = friend_count, data = pf, binwidth = 0.01,
          xlab = 'Number of friends',
          ylab = 'Number of Facebook users',
          color = I('red'), fill = I('#F79420')) +
scale_x_log10()

p3 = qplot(x = friend_count, data = pf, 
          xlab = 'Number of friends',
          ylab = 'Number of Facebook users',
          color = I('red'), fill = I('#F79420')) +
scale_x_sqrt()

grid.arrange(p0, p1, p2, p3, ncol=2)
# check and save!
```

```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3)
```

***


### Frequency Polygons
Notes: Frequency polygon is used to present two or more distributions on one plot for comparison

Looking at:

    qplot(x = friend_count, data = pf, binwidth = 10) +
      scale_x_continuous(limits = c(0, 1000),
                         breaks = seq(0, 1000, 50)) +
      facet_wrap(~gender)
      
By default qplot creates frequency histogram when passed one variable. Use geom parameter to tell it to do something else, e.g. geom = 'freqpoly' to generate fequency poligon.

The layer `facet_wrap(~gender)` can be replaced by parameter `color = gender` to have both lines on a single plot.

To show proportions instead of counts add y axes: `y = ..count../sum(..count..)`


```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10,
      geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
#names(pf)
qplot(x = www_likes, data = subset(pf, !is.na(gender)), binwidth = 1,
      geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0,100), breaks = seq(0,100,10)) +
  scale_y_log10()
```
```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)), 
      geom = 'freqpoly', color = gender) +
  scale_x_continuous() +
  scale_x_log10()
```

Who has more likes, males or females? Quick answer.
```{r}
by(pf$www_likes, pf$gender, sum)
table(pf$gender)
```

***

### Box Plots
Notes: The interquartile range or IQR includes all of the values between the bottom and top of the boxes in the boxplot (http://en.wikipedia.org/wiki/Interquartile_range). 

[Visualization] (http://en.wikipedia.org/wiki/File:Boxplot_vs_PDF.svg) of the IQR with a normal probability distribution function with $\mu=1$ and $\sigma^2=1$ (pdf). 

```{r Box Plots}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot')

```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0,1000))
```
Limiting the axes without discarding data points
```{r}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0,1000))
```
***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
by(pf$friendships_initiated, pf$gender, summary)


```

Response:

***

### Getting Logical
Notes: Binary transformation

```{r Getting Logical}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

summary(pf$mobile_check_in) / sum(summary(pf$mobile_check_in))

sum(pf$mobile_check_in == 1) / length(pf$mobile_check_in)

```

Response:

***

### Analyzing One Variable
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!
